---
title: "Data Science for Public Policy"
author: "Xiyu Zhang - xz551"
output:
  pdf_document: default
  html_document:
    df_print: paged
urlcolor: blue
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(warning = FALSE)

```

```{r include = FALSE}
def.chunk.hook  <- knitr::knit_hooks$get("chunk")
knitr::knit_hooks$set(chunk = function(x, options) {
  x <- def.chunk.hook(x, options)
  ifelse(options$size != "normalsize", paste0("\n \\", options$size,"\n\n", x, "\n\n \\normalsize"), x)
})

```

\begin{center}
\Huge{PPOL 670 | Assignment 3}

\Huge{Applied Introduction to R's Tidyverse}
\end{center}

\vspace{0.1in}

```{r}
library(tidyverse)
library(ipumsr)
library(srvyr)
library(haven)

```


# Exercise 01 (1 point)

```{r}
library(ipumsr)
library(tidyverse)
ddi <- read_ipums_ddi("cps_00002.xml")
data <- read_ipums_micro(ddi)
glimpse(data)
```


\newpage

# Exercise 02 (2 points)
#2.
```{r}
library(srvyr)
cps_svy <- as_survey_design(data, weights = WTFINL)
class(data)
class(cps_svy)

```
#3. The difference is that, the as_survey_design function turn the tibble into a "tbl_svy", object, and create two new catagories called "survey.design" & "survey.design2"

#4. Yes, there is a value of NIU: 
#for UHRSWORKT, 1999-onward ASEC:0 = No hours; 997 = Hours vary; 999 = Not in universe (NIU)
#for AHRSWORKT, 999 = NIU (Not in universe)

```{r}
NIU.UHR.AHR <- data %>%
  group_by(YEAR, UHRSWORKT, AHRSWORKT) %>%
  summarise(
    unweighted = n()
  ) %>%
  filter(UHRSWORKT == 999 | AHRSWORKT == 999) %>%
  arrange(YEAR, desc(UHRSWORKT))
NIU.UHR.AHR
```

#observed that for all observations in 2020 and 2021, while UHRSWORKT == 999, then AHRSWORKT is inevitabily 999, that means to sum up the observations that are not in the universe in these two years, we could only group_by the AHRSWORKT variable. Rewrite our code:

```{r}
NIU.UHR.AHR <- data %>%
  group_by(YEAR, AHRSWORKT) %>%
  summarise(
    unweighted = n()
  ) %>%
  filter(AHRSWORKT == 999) %>%
  arrange(YEAR, desc(AHRSWORKT))
NIU.UHR.AHR
```

#Thus, in 2020, there were 63321 observations that are not in the universe for  UHRSWORKT and AHRSWORKT; in 2021, there were 61703 of them.

#5.Filter out observations not in the universe for UHRSWORKT and create a new data frame called cps_subset_svy with the assignment operator.

```{r}
cps_subset_svy <- data %>%
  filter(UHRSWORKT != 999)
cps_subset_svy
```

#6.Count the number of unweighted responses for each value for UHRSWORKT.
```{r}
cps_subset_6 <- cps_subset_svy %>%
  group_by(UHRSWORKT) %>%
  filter(UHRSWORKT != 997) %>%
  summarize (
    unweighted = n()
  ) 
ggplot(data = cps_subset_6) + 
  geom_col(mapping = aes(x = UHRSWORKT, y = unweighted))
```

#7.Calculate the mean usual hours worked (UHRSWORKT) in 2020 and 2021. Exclude workers with "Hours vary".
```{r}
cps_subset_7 <- cps_subset_svy %>%
  group_by(YEAR) %>%
  filter(UHRSWORKT != 997) %>%
  summarise(mean_UHRSWORKT = mean(UHRSWORKT))
cps_subset_7
```

#8.Calculate the proportion of workers who usually worked exactly 40 hours in 2021.
```{r}
cps_subset_8 <- cps_subset_svy %>%
  filter(YEAR == 2021) %>%
  mutate(
    HOURS_40 = if_else(condition = UHRSWORKT == 40, true = 1, false = 0)
    ) %>%
  group_by(YEAR) %>%
  summarise(proportion_40 = mean(HOURS_40))
cps_subset_8
```

#9.Calculate the proportion of workers who worked less, the same, and more than usual in April 2020 and April 2021 (separately) by comparing UHRSWORKT and AHRSWORKT.
```{r}
cps_subset_91 <- cps_subset_svy %>%
  filter(UHRSWORKT != 997) %>%
  select(YEAR, UHRSWORKT, AHRSWORKT) %>%
  mutate(
    AprilWork = case_when(
      AHRSWORKT > UHRSWORKT ~ "work more",
      AHRSWORKT < UHRSWORKT ~ "work less",
      TRUE ~ "work the same"
    )
  ) 
cps_subset_92 <- as_survey_design(cps_subset_91) %>%
  group_by(YEAR, AprilWork) %>%
  summarise(prop = survey_prop())
cps_subset_92
```
\newpage

# Exercise 03 (2 points)
#2. Using cps_svy, calculate the weighted count of observations ages 16 or older for each year with survey_count(). Your results should match the April 2020 and April 2021 values to the closest thousand because of rounding.
#Sorry I don't have any idea about "survey_count", it seems not appear when I was trying to search it. I also don't know how to put the age >= 16 into weight.

#3. Create three numeric indicator variables.
```{r}
cps_svy_33 <- cps_svy %>%
  mutate(
    labor_force = case_when(
      LABFORCE == 2 ~ 1,
      LABFORCE == 1 ~ 0,
      TRUE ~ 999
    )
  ) %>%
  mutate(
    employed = case_when(
      EMPSTAT == 10 | EMPSTAT == 12 ~ 1,
      EMPSTAT == 20 | EMPSTAT == 21 | EMPSTAT == 22 ~ 0,
      TRUE ~ 999
    )
  ) %>%
  mutate(
    unemployed = case_when(
      EMPSTAT == 10 | EMPSTAT == 12 ~ 0,
      EMPSTAT == 20 | EMPSTAT == 21 | EMPSTAT == 22 ~ 1,
      TRUE ~ 999
    )
  )
```

#4.Filter to the civilian population ages 16 or older, and calculate the relavant population.
```{r}
cps_svy_34 <- cps_svy_33 %>%
  filter(AGE >= 16) %>%
  filter(POPSTAT == 1) %>%
  summarise(
    LABFOR_cal = survey_total(labor_force == 1),
    EMPLOY_cal = survey_total(employed == 1),
    UNEMPLOY_cal = survey_total(unemployed == 1)
            )
cps_svy_34
```

#5. Compare the result with microdata to the official tabulation for April 2021.
#The official data is: civilian labor force - 160,379,000; employed - 151,160,000; unemployed - 9,220,000.
#What I got here are (shown above): civilian labor force - 316,601,435; employed - 284,601,546; unemployed - 31,999,889. They are far from "close", thus there might be some problems in my code.

\newpage

# Exercise 04 (2 points)
#1.-3
```{r}
cps2021 <- filter(data, YEAR == 2021)
cps2021_sum <- cps2021 %>%
  summarise(n())
cps2021_sum
nber <- read_dta("cpsb202104.dta")
nber_sum <- nber %>%
  select(hufinal, pwcmpwgt) %>%
  summarise(n())
nber_sum
```
#3. Notice that n = 133,449. There are 133,449 observations in dataset "nber", while there are 111,003 observations in dataset "cps2021".

#4. As it shows, there are 111,003 observations in nber_44, which match with the amount of the observations in dataset "cps2021".
```{r}
nber_44 <- filter(nber, hufinal <= 205)
summarise(nber_44, n())
```

#5. Combine the columns.
```{r}
nber_cps <- bind_cols(cps2021, nber_44)
nber_cps
```

#6.-9. adjust the weights to previous results.
```{r}
nber_cps_svy <- nber_cps %>%
  mutate(pwcm_cal = pwcmpwgt/10000) %>%
  as_survey_design(weights = pwcm_cal) %>%
  mutate(
    labor_force = case_when(
      LABFORCE == 2 ~ 1,
      LABFORCE == 1 ~ 0,
      TRUE ~ 999
    )
  ) %>%
  mutate(
    employed = case_when(
      EMPSTAT == 10 | EMPSTAT == 12 ~ 1,
      EMPSTAT == 20 | EMPSTAT == 21 | EMPSTAT == 22 ~ 0,
      TRUE ~ 999
    )
  ) %>%
  mutate(
    unemployed = case_when(
      EMPSTAT == 10 | EMPSTAT == 12 ~ 0,
      EMPSTAT == 20 | EMPSTAT == 21 | EMPSTAT == 22 ~ 1,
      TRUE ~ 999
    )
  ) %>%
  filter(AGE >= 16) %>%
  filter(POPSTAT == 1) %>%
  summarise(
    LABFOR_cal = survey_total(labor_force == 1),
    EMPLOY_cal = survey_total(employed == 1),
    UNEMPLOY_cal = survey_total(unemployed == 1)
            )
nber_cps_svy
```

#10.The adjust results came out to be:labor force = 160,379,460; employed = 151,159,727; unemployed = 9,219,734.
#Compared with the official data: civilian labor force = 160,379,000; employed = 151,160,000; unemployed = 9,220,000.
#Now my results are very close to the official one.
\newpage

# Exercise 05 (2 points)
```{r}
library(ggplot2)
p <- ggplot() +
  geom_bar(data = cps_subset_92, 
           aes(x = AprilWork, y = prop, fill = YEAR), 
           stat = "identity",
           position = position_dodge())
p + ggtitle(label = "Working Hours 2020 & 2021",
  subtitle = "the proportion of workers comparing working hours between April 2020 and April 2021") +
  labs(caption = "Data source: CPS") + 
  xlab("April Work (type)") + ylab("proportion")
```
